---
title: "HPC rrBLUP script"
author: "Joseph Oddy"
date: '2022-06-14'
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```
Load rrBLUP and dplyr packages.

```{r}
library(rrBLUP)
library(dplyr)
```

Load genotype and phenotype files and impute missing markers.

```{r}
# import geno and pheno files
H18.geno <- as.matrix(read.csv(file = "RXC_rrBLUP_genotypes.csv", header = TRUE))
rownames(H18.geno) <- H18.geno[,1]
H18.geno <- H18.geno[,-1]

H18.pheno <- as.matrix(read.csv(file = "220124 H18 data.csv", header = TRUE))
rownames(H18.pheno) <- H18.pheno[,1]
H18.pheno <- H18.pheno[,-1]

H19.pheno <- as.matrix(read.csv(file = "220124 H19 data.csv", header = TRUE))
rownames(H19.pheno) <- H19.pheno[,1]
H19.pheno <- H19.pheno[,-1]

# impute missing markers
impute <- A.mat(H18.geno,max.missing = 0.5,impute.method="mean",return.imputed=T)
Markers_impute <- impute$imputed
```

Create function for estimating marker effects with five-fold cross validation.

```{r}
# creat empty matrix for results
res<-matrix(NA,5,2)

### function for getting testing and training sets
fivefoldcv <- function(geno, pheno, ...){
  
  for (i in 1:5){
  # create vector to assign rows to 5 different groups
  linegrouping <-  sample(rep(1:5, length.out = 171))
  linegrouping <- factor(linegrouping)
  assigngroups <- data.frame(rownames(geno), linegrouping)
  test <- assigngroups[which(linegrouping == i),]
  test <- test[,1]
  train <- subset(assigngroups, linegrouping != i)
  train <- train[,1]
  
  # get pheno and geno data for the training and testing
  pheno_train <- pheno[train,]
  geno_train <- Markers_impute[train,]
  pheno_test <- pheno[test,]
  geno_test <- Markers_impute[test,]
  trait <- pheno_train
  traitname <- colnames(pheno)

  # calculate marker effects
  mxslv <- mixed.solve(trait, Z = geno_train, K = NULL, SE = FALSE, return.Hinv=FALSE)
  TPr <- mxslv$u
  effects <- as.matrix(TPr)
  pred_test <- geno_test %*% effects
  predictions <- (pred_test[,1]) + mxslv$beta
  
  # calculate correlation between test predictions and validation set
  validation <- pheno_test
  accuracy <- cor(predictions, validation, use = "complete")
  
  # Store the results
  res[i,1] <- accuracy
  res[i,2] <- traitname
  
  }
  return(data.frame(res))
}
```

Create and run functions for looping the "fivefoldcv" function over the columns of both H18 and H19 datasets. Code for running within year prediction estimates is shown here.

```{r}
#### new function for looping over columns H18

list1 <- list()
traitlooping <- function(geno, pheno, ...){
  for (i in 1:ncol(H18.pheno)){
    pheno <- H18.pheno[,i, drop = FALSE]
    list1[[i]] <- fivefoldcv(geno, pheno)
  }
  return(list1)
}

### repeat 10000 times H18
list2 <- list()
suppressWarnings(for (i in 1:10000){
  H18 <- traitlooping(H18.geno, H18.pheno)
  list2[[i]] <- H18
})

#### new function for looping over columns H19

list3 <- list()
traitlooping2 <- function(geno, pheno, ...){
  for (i in 1:ncol(H19.pheno)){
    pheno <- H19.pheno[,i, drop = FALSE]
    list3[[i]] <- fivefoldcv(geno, pheno)
  }
  return(list3)
}

### repeat 10000 times H19
list4 <- list()
suppressWarnings(for (i in 1:10000){
  H19 <- traitlooping2(H18.geno, H19.pheno)
  list4[[i]] <- H19
})
```

Create and run function for formatting the output.

```{r}
### function for formatting genomic prediction output
formatting <- function(gpdata){
  
  # unlist
  unlisted <- unlist(gpdata, recursive = FALSE)
  
  # bind results into dataframe and order
  genpredicted <- do.call(rbind.data.frame, unlisted)
  genpredicted <- genpredicted[order(genpredicted$X2),]
  
  # calculate means over permutations
  genpredicted$X1 <- as.numeric(genpredicted$X1)
  genpredicted$X2 <- as.factor(genpredicted$X2)
  names(genpredicted)[1] <- "Correlation"
  names(genpredicted)[2] <- "Trait"
  genmeans <- aggregate(genpredicted$Correlation~genpredicted$Trait, FUN=mean)
  names(genmeans)[1] <- "Trait"
  names(genmeans)[2] <- "Correlation"
  genmeans <- genmeans[order(-genmeans$Correlation),]
  
  return(list(genpredicted, genmeans))
}

### formatting
H18.summary <- formatting(list2)

H19.summary <- formatting(list4)

write.csv(H18.summary[[1]], file = "H18 summary")
write.csv(H19.summary[[1]], file = "H19 summary")
```

